---
title: "Which DVS tribe do you belong?"
output: html_notebook
---

The DVS community is a diverse community made up of people from different backgrounds, demographics and experiences. Have you ever wondered what could be the different possible profiles that are made up of this community and which of these profiles do you identify yourself the closest to?

Let's look at the 4 different indicators and mix-match them to create different profiles of what the DVS population are made out of:

* **Years of experience** (Junior vs Senior)
* **Role** (Creative vs Analytical)
* **Income group** (Humbly-paid vs Well-paid)
* **Commitment** (Hobbyist vs Professional)

#### Percentage distribution of tribes according to different combination of indicators
```{r}
library("tidyverse")
source("scripts/clean.R")

tribe_tally

tribe_plot <- tribe_tally %>% 
         mutate(group = paste(Experience, RoleType, IncomeGroup, Commitment)) %>%
         mutate(group = fct_reorder(group, n))

ggplot(tribe_plot, aes(reorder(group, nPerc), nPerc)) + 
  geom_col(fill = "blue") +
  coord_flip()
  
```


## Years of experience (Junior vs Senior)

Median years of experience is 6 years. You are considered a **junior** if your DV experience is less than 6 years, while **senior** if more than 6 years.
```{r}
experience_sorted

ggplot(experience_sorted, aes(YearsDVExperience, weight = n, fill = experience)) + 
  geom_bar() +
  coord_flip()

```

## Role (Creative vs Analytical)

Around 44% of the DVS members have analytical-leaning roles while only 14% belong to the creative-leaning roles. Meanwhile, some roles have been categorised as ambiguous and could hardly be categorized such as leadership and cartography. They comprise 18% of the DVS population. 22% chose to not indicate their roles.

While roles are usually requires to be both creative or analytical in function and purpose, we attempt to classify roles that are more likely leaning towards either of them.

### Creative
Roles that lean more towards creative or humanities related subjects:

* Designer
* Journalist
* Teacher
* Cartographer

### Analytical
Roles that lean more toward analytical or rational related subjects:

* Analyst
* Developer
* Scientist
* Engineer
* Leadership (Manager, Director, VP, etc.)


#### 44% are Analytical and less than half of it are Creatives
```{r} 
role_tally

```

```{r, fig.width=10, fig.height=1.5} 
ggplot(role_tally %>% filter(RoleType != "unknown"), 
       aes(RoleType, weight = nPerc, fill = RoleType)) +
  geom_bar() +
  coord_flip()

```
## Income group (Humbly-paid vs Well-paid)

Median annual income is at $81,300. You are considered **humbly-paid** if your'e annual income falls below the median annual income and **well-paid** if above it.
```{r}
income_sorted

ggplot(income_sorted, aes(PayAnnual, weight = n, fill = income_group)) +
  geom_bar() +
  coord_flip()
```

## Commitment (Hobbyist vs Employee)

### Hobbyist

Hobbyists are those who do datavis driven with passion and perseverance. They're group together who lean towards:

* Worked on a non-paid project 
* Worked as a freelancer / consultant 
* A student 
* Earning passive income from datavis projects 

### Employee
Professionals are those who are lean towards working as an employee or working in the academe as a teacher.

#### Near 2/3 are professionals while more than 1/3 are hobbyists
```{r}
commitment_tally
```
#### Hobbyists vs Professionals (and Ambiguous) distribution
```{r, fig.width=10, fig.height=2}
ggplot(commitment_tally, aes(Commitment, weight = nPerc, fill = Commitment)) +
  geom_bar() + 
  coord_flip()
```

## Additional indicators - Tools for DV, Charts used and Frustrations
```{r}
source("scripts/add_indicators.R")
tribe_tools
tribe_charts
tribe_frustrations
```

### Tool preferences per tribe (Top 10 tools overall)

```{r, fig.width=12, fig.height=10}
tribe_tools
tools_top10

ggplot(tribe_tools %>% 
         filter(tools_for_dv_vars %in% tools_top10), 
       aes(tools_for_dv_names, percN, fill = tools_for_dv_type)) +
  geom_col() +
  coord_flip() +
  facet_wrap(vars(tribe_desc))
```
#### Observations
* Among the junior-analyticals, the humblypaids' use excel as top tool while the wellpaids' use Tableau. Does this imply that Tableau is for the well-offs? Do they
* Humblypaid creatives proudly shows their affinity for illustrator and it's just so obvious with the charts.
* Unanimously, any tribe seems to use Excel as their top tool for DV.
* Junior creative hobbyists tend to use R as either 2nd or 3rd tool.

```{r, fig.width = 10, fig.height = 10}
# ggplot(tribe_tools %>% 
#          ungroup() %>%
#          group_by(tribe_desc) %>%
#          top_n(5, n) %>% ungroup(), aes(tools_for_dv_vars, percN, color = tribe_desc)) +
#   geom_point() +
#   coord_flip()
```


### How endemic tribes are per industry?
```{r, fig.width=12, fig.height=10}
tribe_orgsec

ggplot(tribe_orgsec, aes(org_sector_names, n, fill = org_sector_category)) +
  geom_col() +
  coord_flip() +
  facet_wrap(vars(tribe_desc))
```


### Top frustrations per tribe
```{r, fig.width=12, fig.height=10}
tribe_frustrations

ggplot(tribe_frustrations %>%
         filter(top_frustrations_vars %in% frustrations_top10), 
       aes(top_frustrations_names, percN, fill = top_frustrations_type)) +
  geom_col() +
  coord_flip() +
  facet_wrap(vars(tribe_desc))
```
#### Observations
* Junior-analyticals looks unanimous in their frustrations, the lask of design expertise. For junior-analytical-wellpaid-hobbyist, it is their second frustration to lack of time.
* Junior-creatives' top frustration seems to be the lack of time. It is third for humblypaid-employees and wellpaid-hobbyists where their top frustrations is in the lack of technical skills.
* Junior-creative-wellpaid-hobbyists, which is the rarest tribe, mention that a lack of tech skillw (1st) and lacking mentorship (2nd) are their main frustrations.
* Senior analyticals are unaninous with the lack of time as their top frustrations along with creatives humblypaid-hobbyist and wellpaid-employee.
* Senior-creative-humbly-paid employees notes that the lack of recpect and tech limitations are their top frustrations.


### Top issues per tribe
```{r, fig.width=12, fig.height=10}
tribe_issues

ggplot(tribe_issues %>%
         filter(top_issues_vars %in% top_issues_top10), 
       aes(top_issues_names, percN, fill = top_issues_type)) +
  geom_col() +
  coord_flip() +
  facet_wrap(vars(tribe_desc))
```
#### Observations
* Unanimously lack of literacy for majority regardless of tribe. Second to it is the lack of awareness of impact of data visualization
* If we try to remove the top 2 



